SubsidyExplorer - Step 2: Create Vessel List

### 2020-03-26 scripts/02_vessel_list.Rmd
more vessel list work

This script creates the global vessel list used in the SubsidyExplorer toolkit. It requires access to the raw Global Fishing Watch (GFW) data stored on BigQuery (not-publically available), as well as a valid Google Cloud Console billing account. It outputs a single data file to `r here::here("results", "02-vessel-list")`:

- vessel_list_2018_raw.csv

It also creates a 1x1 degree version for use in making higher resolution figures. This is output to

@@

# Path to cleaned landings data file
#catch_dat_path <- here::here("data", "edited", "fisheries-area-2000-2017-tidy.csv")

@@

```{r}
# Code options
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = F)

# Directories for the output files generated by this script
results_dir <- here::here("results", "02-vessel-list/")
if (dir.exists(results_dir) == F) {
  dir.create(results_dir, recursive = T)
}

@@

We first extract the number of hours spent fishing by each of our good fishing vessels in each 5x5 degree latitude/longitude bin from the raw fishing effort table (`world-fishing-827.gfw_research.pipe_v20190502_fishing`).

@@

We then extract a vessel table with the the following attributes: 
- year (int): Year. Matches our data year specified above.
- lat_bin_center (int): 1 degree latitude bin, coordinates refer to the center of the bin
- lon_bin_center (int): 1 degree longitude bin, coordinates refer to the center of the bin
- fao_region (int): Major FAO statistical region
- fishing_hours_lat_lon: Total annual fishing hours in the specified lat/lon bin by a single vessel
- ssvid (chr): Unique vessel identification number given by GFW (MMSI number for most vessels)
- flag (chr): Best flag state of the vessel
- engine_power_kw (int): Best engine power of the vessel

@@

binned_effort_table <- bq_table(project = bq_project, 
                                table = "effort_5deg_good_fishing_vessels_2018_raw",
                                dataset = "SubsidyExplorer")

@@

  bq_table_delete(binned_effort_table)
  effort_binned_good_fishing_vessels_2018 <- 
    bq_project_query(bq_project, 
                    sql2, 
                    destination_table = binned_effort_table, 
                    use_legacy_sql = FALSE, 
                    allowLargeResults = TRUE) %>%
    bq_table_download()

}else {
  bq_table_create(binned_effort_table)
  effort_binned_good_fishing_vessels_2018 <- 
    bq_project_query(bq_project, 
                   query = sql2, 
                   destination_table = binned_effort_table,
                   use_legacy_sql = FALSE, 
                   allowLargeResults = TRUE) %>%
    bq_table_download()
}

# Save
write_csv(effort_binned_good_fishing_vessels_2018, paste0(results_dir, "effort_binned_good_fishing_vessels_2018_raw.csv"))
```

@@

We then extract a vessel table with the the following attributes: 
- year (int): Year. Matches our data year specified above.
- eez_id (chr): EEZ identifier. 
- fao_region (int): Major FAO statistical region
- fishing_hours_eez_fao: Total annual fishing hours in the specified eez and/or fao region by a single vessel
- ssvid (chr): Unique vessel identification number given by GFW (MMSI number for most vessels)
- flag (chr): Best flag state of the vessel. 
- vessel_class (chr): Best vessel type. 
- length_m (dbl): Best vessel length in meters.
- tonnage_gt (dbl): Best vessel tonnage in gross tons. 
- engine_power_kw (dbl): Best engine power in killowatt hours. 
- broadcast_shipname (chr): Most common ship name broadcast by the vessel.
- broadcast_callsign (chr): Most common ship callsign broadcast by the vessel.
- broadcast_imo (chr): Most common imo number broadcast by the vessel.
- territory1_iso3 (chr) : Three letter code cooresponding to the administering territory of the EEZ
- territory2_iso3 (chr) : Three letter code cooresponding to the second administering territory of the EEZ (disputed areas only)
- territory3_iso3 (chr) : Three letter code cooresponding to the third administering territory of the EEZ (disputed areas only)
- sovereign1_iso3 (chr) : Three letter code cooresponding to the sovereign state of the EEZ
- sovereign2_iso3 (chr) : Three letter code cooresponding to the second sovereign state of the EEZ (disputed areas only)
- sovereign3_iso3 (chr) : Three letter code cooresponding to the third sovereign state of the EEZ (disputed areas only)

@@

  # This subquery identifies good fishing vessels from our previous query
  good_vessels AS (
  SELECT
    *
  FROM
    `ucsb-gfw.SubsidyExplorer.all_good_fishing_vessels`),

@@

SELECT
  *,
  fishing_hours_eez_fao*engine_power_kw as fishing_KWh_eez_fao
FROM
  fishing_by_eez_and_fao
LEFT JOIN
  vessel_info
USING
  (ssvid,
  year)
LEFT JOIN
  eez_info
USING
  (eez_id)
'

effort_table <- bq_table(project = bq_project, 
                     table = "effort_good_fishing_vessels_2018_raw",
                     dataset = "SubsidyExplorer")

@@

  bq_table_delete(effort_table)
  effort_good_fishing_vessels_2018 <- 
    bq_project_query(bq_project, 
                    sql3, 
                    destination_table = effort_table, 
                    use_legacy_sql = FALSE, 
                    allowLargeResults = TRUE) %>%
     bq_table_download()

}else {
  bq_table_create(effort_table)
  effort_good_fishing_vessels_2018 <- 
    bq_project_query("world-fishing-827", 
                   query = sql3, 
                   destination_table = effort_table,
                   use_legacy_sql = FALSE, 
                   allowLargeResults = TRUE) %>%
    bq_table_download()
}

# Save
write_csv(effort_good_fishing_vessels_2018, paste0(results_dir, "effort_good_fishing_vessels_2018_raw.csv"))
```

@@

```{r}
# EU countries
#Belgium, Bulgaria, Croatia, Cyprus, Denmark, Estonia, Finland, France, Germany, Greece, Ireland, Italy, Latvia, Lithuania, Malta, Netherlands, Poland, Portugal, Romania, Slovenia, Spain, Sweden, United Kingdom
eu_countries <- c("BEL", "BGR", "HRV", "CYP", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "IRL", "ITA", "LVA", "LTU", "MLT", "NLD", "POL", "PRT", "ROU", "SVN", "ESP", "SWE", "GBR")

# EU overseas territories: 
# Dutch overseas territories: Aruba, Bonaire/St. Eustatius/Saba, Curacao, Sint Maarten
dutch_territories <- c("ABW", "BES", "CUW", "SXM")

# French overseas territories: French Guiana, French Polynesia, French Southern and Antarctic Territories, Guadeloupe, Martinique, Mayotte, New Caledonia, Reunion, Saint Barthelemy, Saint Martin, St. Pierre and Miquelon, Wallis and Futuna Islands
french_territories <- c("GUF", "PYF", "ATF", "GLP", "MTQ", "MYT", "NCL", "REU", "BLM", "MAF", "SPM", "WLF")

# Danish overseas territories: Faroe Islands, Greenland
danish_territories <- c("FRO", "GRL")

# British overseas territories: Anguilla, Ascension island, Bermuda, Cayman Islands, Falkland Islands, Guernsey, Gibraltar, British Indian Ocean Territories, Jersey, Monserratt, Pitcairn Island, South Georgia and the South Sandwich Islands, Saint Helena/Ascension/Tristan da Cunha, Tristan da Cunha, Turks and Caicos Islands, British Virgin Islands
uk_territories <- c("AIA","ASC","BMU","CYM","FLK","GGY","GIB","IOT","JEY","MSR","PCN","SGS","SHN","TAA","TCA","VGB")

# Combine
eu_territories <- c(dutch_territories, french_territories, danish_territories, uk_territories)

# Norwegian overseas territories: Bouvet Islands, Svalbard and Jan Mayen Islands
norwegian_territories <- c("BVT", "SJM")

# US overseas territories:  American Samoa, Guam, Northern Marianas Islands, Puerto Rico, US Virgin Islands
us_territories <- c("ASM", "GUM", "MNP", "PRI", "VIR")
```

@@

```{r}
# Load catch and revenue data
catch_flag_fao <- read_csv(catch_dat_path) %>%
  dplyr::filter(year == 2017) %>%
  dplyr::filter(!(iso3 %in% c("EU", "USA-T", "UNK"))) %>%
  group_by(iso3, area, variable) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  ungroup() %>%
  spread(variable, value) %>%
  dplyr::filter(landings > 0)

>>>>

@@

# Path to cleaned capture production data file
catch_dat_path <- here::here("data", "edited", "fao_2019_capture_production_tidy.csv")

# Path to cleaned landed value data file
landed_value_dat_path <- here::here("data", "edited", "estimated_landed_value_isscaap_groups_tidy.csv")

@@

# Code options
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = F)

# Directories for the output files generated by this script
results_dir <- here::here("results", "02-vessel-list/")
if (dir.exists(results_dir) == F) {
  dir.create(results_dir, recursive = T)
}
```

This script creates the global vessel list used in the SubsidyExplorer toolkit. It requires access to the raw Global Fishing Watch (GFW) data stored on BigQuery (not-publically available), as well as a valid Google Cloud Console billing account. It creates and outputs the raw 2018 global vessel list (`r paste0(results_dir, "vessel_list_2018_raw.csv")`) with effort by vessel aggregated to the EEZ and/or FAO statistical region.

It also creates a 1x1 degree version for use in making higher resolution figures (`r paste0(results_dir, "vessel_list_2018_1x1_binned.csv")`)

@@

We first extract the number of hours spent fishing by each of our good fishing vessels in each 1x1 degree latitude/longitude bin from the raw fishing effort table (`world-fishing-827.gfw_research.pipe_v20190502_fishing`).

@@

We then extract a vessel table with the the following attributes: 
- year (int): Year. Matches our data year specified above.
- lat_bin_center (int): 1 degree latitude bin, coordinates refer to the center of the bin
- lon_bin_center (int): 1 degree longitude bin, coordinates refer to the center of the bin
- fao_region (int): Major FAO statistical region
- ssvid (chr): Unique vessel identification number given by GFW (MMSI number for most vessels)
- flag (chr): Best flag state of the vessel
- engine_power_kw (int): Best engine power of the vessel
- fishing_hours_lat_lon: Total annual fishing hours in the specified lat/lon bin by the vessel
- fishing_KWh_lat_lon: Total annual fishing KWh in the specified lat/lon bin by the vessel

@@

binned_effort_table <- bq_table(project = bq_project, 
                                table = "effort_1x1_good_fishing_vessels_2018_raw",
                                dataset = "SubsidyExplorer")

@@

  bq_table_delete(binned_effort_table)
  effort_1x1_good_fishing_vessels_2018 <- 
    bq_project_query(bq_project, 
                    sql2, 
                    destination_table = binned_effort_table, 
                    use_legacy_sql = FALSE, 
                    allowLargeResults = TRUE) %>%
    bq_table_download()

}else {
  bq_table_create(binned_effort_table)
  effort_1x1_good_fishing_vessels_2018 <- 
    bq_project_query(bq_project, 
                   query = sql2, 
                   destination_table = binned_effort_table,
                   use_legacy_sql = FALSE, 
                   allowLargeResults = TRUE) %>%
    bq_table_download()
}

# Save
write_csv(effort_1x1_good_fishing_vessels_2018, paste0(results_dir, "effort_1x1_good_fishing_vessels_2018_raw.csv"))
```

@@

We then extract a vessel table with the the following attributes: 
- year (int): Year. Matches our data year specified above.
- eez_id (chr): EEZ identifier. 
- fao_region (int): Major FAO statistical region
- ssvid (chr): Unique vessel identification number given by GFW (MMSI number for most vessels)
- flag (chr): Best flag state of the vessel. 
- vessel_class (chr): Best vessel type. 
- length_m (dbl): Best vessel length in meters.
- tonnage_gt (dbl): Best vessel tonnage in gross tons. 
- engine_power_kw (dbl): Best engine power in killowatt hours. 
- broadcast_shipname (chr): Most common ship name broadcast by the vessel.
- broadcast_callsign (chr): Most common ship callsign broadcast by the vessel.
- broadcast_imo (chr): Most common imo number broadcast by the vessel.
- territory1_iso3 (chr) : Three letter code cooresponding to the administering territory of the EEZ
- territory2_iso3 (chr) : Three letter code cooresponding to the second administering territory of the EEZ (disputed and/or joint regime areas only)
- territory3_iso3 (chr) : Three letter code cooresponding to the third administering territory of the EEZ (disputed and/or joint regime areas only)
- sovereign1_iso3 (chr) : Three letter code cooresponding to the sovereign state of the EEZ
- sovereign2_iso3 (chr) : Three letter code cooresponding to the second sovereign state of the EEZ (disputed and/or joint regime areas only)
- sovereign3_iso3 (chr) : Three letter code cooresponding to the third sovereign state of the EEZ (disputed and/or joint regime areas only)
- fishing_hours_eez_fao: Total annual fishing hours in the specified eez and/or FAO statistical region by the vessel
- fishing_KWh_eez_fao: Total annual fishing KWh hours in the specified eez and/or FAO statistical region by the vessel

@@

@@

SELECT
  *,
  CASE
    WHEN engine_power_kw IS NULL THEN fishing_hours_eez_fao*400
    ELSE fishing_hours_eez_fao*engine_power_kw
  END AS fishing_KWh_eez_fao
FROM
  fishing_by_eez_and_fao
LEFT JOIN
  vessel_info
USING
  (ssvid,
  year)
LEFT JOIN
  eez_info
USING
  (eez_id)
'

effort_table <- bq_table(project = bq_project, 
                     table = "effort_eez_fao_good_fishing_vessels_2018_raw",
                     dataset = "SubsidyExplorer")

@@

  bq_table_delete(effort_table)
  effort_good_fishing_vessels_2018 <- 
    bq_project_query(bq_project, 
                    query = sql3, 
                    destination_table = effort_table, 
                    use_legacy_sql = FALSE, 
                    allowLargeResults = TRUE) %>%
     bq_table_download()

}else {
  bq_table_create(effort_table)
  effort_good_fishing_vessels_2018 <- 
    bq_project_query(bq_project, 
                   query = sql3, 
                   destination_table = effort_table,
                   use_legacy_sql = FALSE, 
                   allowLargeResults = TRUE) %>%
    bq_table_download()
}

# Save
write_csv(effort_good_fishing_vessels_2018, paste0(results_dir, "effort_eez_fao_good_fishing_vessels_2018_raw.csv"))
```

@@

```{r}
# Load country dependencies helper file
country_dependencies <- read_csv(here::here("data", "lookup-tables", "country_dependencies.csv"))

# EU countries
eu_countries <- country_dependencies$iso3[country_dependencies$is_EU]

# Dutch overseas territories 
dutch_territories <- country_dependencies$iso3[country_dependencies$is_overseas_territory & country_dependencies$sovereign_iso3 == "NLD"]

# French overseas territories 
french_territories <- country_dependencies$iso3[country_dependencies$is_overseas_territory & country_dependencies$sovereign_iso3 == "FRA"]

# Danish overseas territories
danish_territories <- country_dependencies$iso3[country_dependencies$is_overseas_territory & country_dependencies$sovereign_iso3 == "DNK"]

# UK overseas territories
uk_territories <- country_dependencies$iso3[country_dependencies$is_overseas_territory & country_dependencies$sovereign_iso3 == "GBR"]

# EU overseas territories
eu_territories <- country_dependencies$iso3[country_dependencies$is_overseas_territory & country_dependencies$sovereign_iso3 %in% eu_countries]

# US overseas territories:  American Samoa, Guam, Northern Marianas Islands, Puerto Rico, US Virgin Islands
us_territories <- country_dependencies$iso3[country_dependencies$is_overseas_territory & country_dependencies$sovereign_iso3 == "USA"]

# Norwegian overseas territories: 
norwegian_territories <- country_dependencies$iso3[country_dependencies$is_overseas_territory & country_dependencies$sovereign_iso3 == "NOR"]
```

@@

```{r}
# Load catch data
catch_flag_fao <- read_csv(catch_dat_path) %>%
  dplyr::filter(year == 2017) %>%
  dplyr::filter(!(iso3 %in% c("EU", "USA-T", "UNK"))) %>%
  group_by(iso3, fishing_area, variable) %>% 
  summarize(value = sum(value, na.rm = T)) %>%
  ungroup() %>%
  spread(variable, value)
